---
title: "Lab 15 (FINAL LAB)"
author: "Adam Hayes, Erin Omyer, Richard Park, Jasmine Sanchez"
date: "5/1/2019"
output: html_document
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```



###How does state income impact state education overall?
Our team, the Alge-Bros and Inte-Girls, was recently hired by the Secretary of Education, who will be our domain expert for this project. The Secretary of Education has asked us to analyze how state income affects education overall. 


###Importance
This is an important question to analyze because it will be a good indicator for the Secretary of Education to see how effective money is being used for education at the state level.

###Our Data
https://www.kaggle.com/noriuk/us-education-datasets-unification-project

The data our team decided to use came from the kaggle site posted above. We chose this dataset primarily because it gives insightful and relevant data with regards to nationwide education. More importantly, it serves as an economic underpinning so we can analyze different statistics about how effectively money is being spent on education. This dataset will have extensive value because it breaks each component down to the state-level, allowing us to go more in depth with state incomes instead of a national income.

###Our Team's Findings

```{r}
library(tidyverse)
library(modelr)
options(na.action = na.warn)


education <- read.csv("states_all.csv") %>%
  select(-PRIMARY_KEY)%>%
  group_by(YEAR)
education <- education[!is.na(education$ENROLL),]%>%
  mutate(meanEnroll = mean(ENROLL),
         meanRevenue = mean(TOTAL_REVENUE))
mod <- lm(meanEnroll~YEAR*meanRevenue, data = education)
grid <- education %>%
  data_grid(meanRevenue)
grid <- grid %>%
  add_predictions(mod)

ggplot(education, aes(x=YEAR, y = meanRevenue))+
  geom_line()+
  labs(title = "Average Revenue for all States by Year")

ggplot(education, aes(YEAR))+
  geom_point(aes(y = meanEnroll))+
  geom_line(aes(y = pred),data = grid, size = 1)+
  ggtitle("Model for Average # of Enrolled Students vs Yearly Revenue,\ncompared with Actual Values for all states by Year")
  

```

###Conclusion
First and foremost, our team discovered that overall, enrollment numbers for kids younger than college age tend to increase with the increase of state revenue.  We found that the states that consistently have lower revenues tend to have a much harder time maintaining their enrollment numbers.  We also discovered that the mean expenditure used on instruction for education has been steadily increasing over the years. This means that on average, states nationwide have slowly been spending more money on teachers and other instructors, which in turn, has lead to an increase in student enrollment (which was expected). 

###Recommendations
Our findings suggest that we should focus our government resources towards states with consitently lower revenue, because these are the states struggling to maintain enrollment numbers year to year.  We must try to find a way to help states with less revenue spend their money more efficiently on education, so that their enrollment numbers will continue to increase, rather than decrease. Our findings about state expenditures on education can be useful to the Secretary of Education because the more we pay/spend on teachers, the more students are expected to be enrolled in school.




###Adam's Findings

```{r}
dataAdamAll <- education %>%
  group_by(STATE)%>%
  filter( STATE == "CALIFORNIA" || STATE == "TEXAS"|| STATE == "NEW_YORK" || STATE == "FLORIDA" || STATE == "SOUTH_DAKOTA"||STATE == "NORTH_DAKOTA"||STATE == "WYOMING")


dataAdamTOP <- education %>%
  group_by(STATE)%>%
  filter( STATE == "CALIFORNIA" || STATE == "TEXAS"|| STATE == "NEW_YORK" || STATE == "FLORIDA")
ggplot(dataAdamTOP, aes(x = YEAR, y = TOTAL_REVENUE))+
  geom_line(aes(color = STATE))+
  ggtitle("Total Revenue By State for Highest earning states")

model <- lm(ENROLL ~ YEAR*STATE, data = dataAdamTOP)
grid <- dataAdamTOP %>%
  data_grid(YEAR)
grid <- grid %>%
  add_predictions(model)
ggplot(dataAdamTOP, aes(YEAR, color = STATE)) +
  geom_point(aes(y = ENROLL)) +
  geom_line(aes(y = pred), data = grid, size = 1)+
  ggtitle("Model for Total Enrollment for States with Highest Revenue")

dataAdamBottom <- education %>%
  group_by(STATE)%>%
  filter(STATE == "SOUTH_DAKOTA"||STATE == "NORTH_DAKOTA"||STATE == "WYOMING" )
ggplot(dataAdamBottom, aes(x = YEAR, y = TOTAL_REVENUE))+
  geom_line(aes(color = STATE))+
  ggtitle("Total Revenue By State for Lowest earning states")


model <- lm(ENROLL ~ YEAR*STATE, data = dataAdamBottom)
grid2 <- dataAdamBottom %>%
  data_grid(YEAR)
grid2 <- grid2 %>%
  add_predictions(model)
ggplot(dataAdamBottom, aes(YEAR, color = STATE)) +
  geom_point(aes(y = ENROLL)) +
  geom_line(aes(y = pred), data = grid2, size = 1)+
  ggtitle("Model for Total Enrollment for States with Lowest Revenue")

```

####Comments
The first two graphs represent the four states that consistently ranked top 4 in terms of state revenue from 1993 to 2016.  Similarly, the second two graphs represent the three states that ranked bottom 3.  What we can see from the graphs representing total state revenue over the years is that every state consistently increased their revenue from 1993 to 2016.  The difference then comes in the two modeled graphs, where we can see the graph representing the wealthiest states have strong positive trends of number of kids enrolled in school increasing, as revenue increases over the years.  The one outlier would be New York, where the number of enrolled kids has actually decreased.  On the other hand, the graph modeling the least wealthy states shows weak linear trends, but they tend to trend negative.  What this is showing is that the states generating much less revenue have a harder time maintaining consistent enrollment numbers.  This graph also shows that until about 2010, the enrollment numbers for these states were consistently decreasing.  However, around 2010 we can see a change in the trend towards increasing enrollment numbers which is a great thing for overall education if we can keep this trend going.

###Erin's Findings

####Does the total revenue for education follow the year? (i.e Does the year matter?)

####Importance: From 1993 to 2016, the amount of money circulating in society has grown drastically. Based on this, it is important to know when tidying and organizing data if specific labels matter.

####New tools used: permutation test

```{r}
educationERIN <- read.csv("states_all.csv") %>%
  select(-PRIMARY_KEY)
education <- education[!is.na(education$TOTAL_REVENUE),]

enrolled <- education[!is.na(education$ENROLL),]

dataErin <- education %>%
  select(YEAR,TOTAL_REVENUE)

median1993_E <- education %>%
  filter(YEAR == 1993) %>%
  summarise(median1993 = median(dataErin$TOTAL_REVENUE))

median2016_E <- education %>%
  filter(YEAR == 2016)%>%
  summarise(median2016 = mean(dataErin$TOTAL_REVENUE))

testgroup <- education %>%
  filter(YEAR == 1993 || YEAR == 2016)

perm_median <- function(perms = 1000, values, n1)
{
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  outputMedian <- c()
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    
    vector1 <- sample(values, n1)
    n2 <- length(values)-n1
    vector2 <- sample(values, n2)
    
    # Compute the sample means for the two groups from 
    # step 2
    
    firstMedian <- median(vector1)
    secondMedian <- median(vector2)
    
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    
    medianDifference <- firstMedian - secondMedian
    
    outputMedian[i] <- medianDifference
  }
  # Return new updated vector, created in step 1
  
  return(outputMedian)
}

revDifference <- median2016_E - median1993_E

results <- perm_median(1000, dataErin$TOTAL_REVENUE, 1000)
resultss <- abs(results)

ggplot()+
  geom_histogram(aes(x=resultss),fill = "pink")+
  geom_vline(xintercept = 200000)+
  stat_bin(binwidth = 300) +
  labs(title = "Distribution of Difference in Medians for 1993 and 2016 in Total Revenue")+
  xlab("Difference")

resultERIN <- sum(resultss<200000)
percentiles_data1 <- resultERIN/1000

```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(revDifference, caption = "The actual difference between 1993 and 2016")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(percentiles_data1, caption = "The percentiles of difference in medians")
```
  
####Comments

From 1993 to 2016, there has been a vast change in the way the education system functions in the United States. Not only have the number of students have increased, the number of teachers and the amount of money fluctuating in the system. We conducted a permutation test on the median total revenue for the year 1993 and 2016. From the graph above, the total revenue between years has tried to remain relatively the same with minimal states with vast variations. Through the permutation test, we found a percentile difference to contain roughly 0.3 percent of the data. This number is a small proportion of the data, which means the years do matter when looking at total revenue from 1993 to 2016.The difference between 1993 and 2016 for total revenue is large due to the various changes society has taken over the recent years. The amount of money fluctuating in the U.S. has increased, leading to more money in the education system to help the success of the students. 


###Richard's Findings

####Does the instruction revenue for education coincide with the year? (i.e Does the year matter?)

####Importance: From 2002 to 2015, money has inflated and there is now drastically more money circulating in society. Based on this, it is important to know when tidying and organizing data if specific labels matter.

####New tools used: two-sample permutation test and quantiles

```{r}
dataRichard2002 <- education %>%
  filter(YEAR == 2002) %>%
  summarise(mean = mean(INSTRUCTION_EXPENDITURE))

dataRichard2015 <- education %>%
  filter(YEAR == 2015) %>%
  summarise(mean = mean(INSTRUCTION_EXPENDITURE))

Difference_In_Expenditure = dataRichard2015 - dataRichard2002
Difference_In_Expenditure

Education_group <- education %>%
  filter(YEAR == 2002 | YEAR == 2015)

#Two-Sample Permutation Test

perm_mean <- function(perms = 1000, values, n1)
{
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  outputMean <- vector("double", perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    
    vector1 <- sample(values, n1)
    vector2 <- setdiff(values, vector1)
    
    # Compute the sample means for the two groups from 
    # step 2
    
    firstMean <- mean(vector1)
    secondMean <- mean(vector2)
    
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    
    meanDifference <- firstMean - secondMean
    
    outputMean[i] <- meanDifference
  }
  # Return new updated vector, created in step 1
  
  return(outputMean)
}

EducationMean <- perm_mean(1000, Education_group$INSTRUCTION_EXPENDITURE, 100)
#realResult <- rnorm(result)
hist(result, col = "green", main = "Difference of Means")

duration = EducationMean
quantile(duration, c(0.05, 0.95))

ggplot()+
  geom_histogram(aes(x=EducationMean), bins = 100,fill = "Red")+
  geom_vline(xintercept = 250000)+
  stat_bin(binwidth = 1000) +
  labs(title = "Distribution of Difference in Means for 2002 and 2015 in Instruction Revenue")+
  xlab("Difference")

Richardeducation <- sum(EducationMean<250000)
percentilesRichard <- Richardeducation/1000
percentilesRichard
```

####Comments
The educational infrastructure in today's society has proven to be dynamic and adaptive. In the United States, there is more money being circulated than ever before, there are more kids being enrolled in school, and there are more ways states around the nation are paying their educational systems. I performed a two-sample permutation test with regards to the mean of instruction expenditure between the two years 2002 and 2015. From my graph above, the instructional revenue between years has slowly increased with minimal states with vast variations. Through the permutation test, I found a percentile difference to contain roughly 31.6 percent of the data. This number is a small proportion of the data, which means the years do matter when looking at instructional revenue from 2002 to 2015.The difference between 2002 and 2015 for total revenue is large due to inflation of the economy. Therefore, in conclusion, the amount of money fluctuating in the states has increased, leading to more money being spent on instructors in the education system which ultimately helps the success of the students (which probably explains why more students are enrolled than ever before). 


###Comparison to Lab 2
Team:

When we first began our journey as Data Scientists, we set the goals of becoming an efficient group and perfect our R studio coding. Through each lab, we have become better each time, learning and exploring new topics in R studio. If we could go back in time, we would tell ourselves that the textbook is efficient in showing examples on methods to code. As well, we would want ot ask more questions when starting out the learning of R studio. 

Richard:

Erin:

Through this course, it has provided a different perspective on the studies of data science as a whole. I came in to the class believing I was going to learn mathematical concepts with a dash of coding. But, it was instantly changed when I learned I had to download a completely different software to code in. While the class is tedious, it has allowed myself to have a different perspective on my future career ideas. If I could go back to the beginning of the semester, I would tell myself to practice more hands on examples outside of the book. While the book is thorough, I think it is easy to get caught up in the same terminology and overlook the overall meaning.

Jasmine:

Adam: Reflectin on my goal from the second lab of the semester, I think I have a much stronger understanding of how to utilize real world data with computer science.  I now am able to take real world data, interperet it, make predictions, and apply my new skills to create findings and conclusions.


###Who Did What?
Adam tidied the data and created a model and graph for the team section, and helped with conclusion and recommendation.  In his individual section he did exploratory data analysis on which states consistently generate the most revenue, and which generate the least.  He then compared these two groups in terms of school enrollment using data modeling, and graphs.